% 03-experimental-setup-and-test-cases.tex
% ---------------------------

% Section Title
\section{EXPERIMENTAL SETUP AND TEST CASES} \label{sec:experimental-setup-and-test-cases}

\subsection{Equipment and Configuration} \label{subsec:equipment-and-configuration}

In this section, we describe the hardware and software configuration used to perform our network performance measurements. Table~\ref{tab:equipment-summary} summarizes the main devices, their interfaces, and relevant specifications. 

\begin{table}[ht]
    \small
    \centering
    \begin{tabular}{@{}l p{0.78\columnwidth}@{}}
    \toprule
    \textbf{Device} & \textbf{Key Specifications} \\
    \midrule
    \textbf{PC1} 
        & \textbf{Victus 16-s1005nl Notebook} \newline
          \textit{Operating System:} Ubuntu 24.04.2~LTS \newline
          \textit{Ethernet Interface:} Realtek RTL8111/8168/8211/8411 \newline
          \textit{Wireless Interface:} Realtek RTL8852BE (802.11ax) \\
    \midrule
    \textbf{PC2} 
        & \textbf{Microsoft Surface Laptop Go~3} \newline
          \textit{Operating System:} Ubuntu 24.10 \newline
          \textit{Ethernet Interface:} via Anker PowerExpand+ USB-C Hub \newline
          \textit{Wireless Interface:} Intel Alder Lake-P CNVi (802.11ax) \\
    \midrule
    \textbf{Router} 
        & \textbf{Vodafone Power Station Wi-Fi~6} \newline
          \textit{Ethernet Ports:} 4 $\times$ 1\,GbE ports \newline
          \textit{Wi-Fi:} Dual-band 802.11ax (2.4\,GHz / 5\,GHz) \\
    \midrule
    \textbf{Cables} 
        & CAT.5E (up to 1\,Gbps) \\
    \bottomrule
    \end{tabular}
    \vspace{0.5cm}
    \caption{Summary of Hardware and Network Configuration}
    \label{tab:equipment-summary}
\end{table}

\vspace{-0.1cm} % TODO: check

This hardware setup allows us to compare Ethernet versus Wi-Fi performance under a consistent router and cabling environment. In the next section, we detail the evaluation scenarios and the measurement methodology.

\subsection{Evaluation Scenarios} \label{subsec:evaluation-scenarios}

We considered three distinct network configurations to assess the performance differences between wired and wireless communications. For each scenario, the theoretical goodput is computed based on the nominal link capacity and protocol efficiency.

\begin{enumerate}
    \item \textbf{Both Ethernet:} \\
    In this scenario, both PC1 and PC2 are connected to the router via CAT.5E cables, providing a nominal link capacity of 1\,Gbps. The efficiency for Ethernet is calculated as follows:
    \[
    \eta_{TCP}^{Eth} \approx \frac{1460}{1460 + 20 + 20 + 38} \approx 94.9\%,
    \]
    \[
    \eta_{UDP}^{Eth} \approx \frac{1472}{1472 + 20 + 8 + 38} \approx 95.7\%.
    \]
    Thus, the expected goodput is:
    \[
    G_{TCP}^{Eth} \leq 0.949 \times 1000\,\text{Mbps} \approx 949\,\text{Mbps},
    \]
    \[
    G_{UDP}^{Eth} \leq 0.957 \times 1000\,\text{Mbps} \approx 957\,\text{Mbps}.
    \]

    \vspace{0.2cm} % TODO: check

    \item \textbf{Both Wi-Fi:} \\
    For this configuration, both devices use their wireless interfaces (802.11ax) to connect to the router. Although the nominal Wi-Fi link speed is assumed to be approximately 867\,Mbps, the half-duplex nature of Wi-Fi effectively halves the throughput available for data transfer. Assuming a Wi-Fi efficiency factor of about 80\%, the expected goodput for TCP is:
    \[
    G_{TCP}^{WiFi} \leq 0.80 \times 867\,\text{Mbps} \times \frac{1}{2} \approx 347\,\text{Mbps},
    \]
    and similarly for UDP:
    \[
    G_{UDP}^{WiFi} \leq 0.806 \times 867\,\text{Mbps} \times \frac{1}{2} \approx 350\,\text{Mbps}.
    \]

    % \vspace{0.1cm}

    \item \textbf{Mixed Scenario:} \\
    In this configuration, one device (PC1) is connected via Ethernet while the other (PC2) uses its Wi-Fi interface. Here, the bottleneck is the Wi-Fi link; however, since only one device is utilizing Wi-Fi, the throughput is not halved. The expected goodput is then:
    \[
    G_{TCP}^{Mixed} \leq 0.80 \times 867\,\text{Mbps} \approx 694\,\text{Mbps},
    \]
    \[
    G_{UDP}^{Mixed} \leq 0.806 \times 867\,\text{Mbps} \approx 698\,\text{Mbps}.
    \]
\end{enumerate}

These calculations provide the theoretical upper bounds for goodput in each scenario. The experimental results, obtained via automated measurements using the provided Python script, are compared against these predictions to evaluate real-world performance.

\medskip

\textbf{Note on Experimental Automation:} \\
The experimental process was automated using a custom Python script. This script manages both the server and client modes of iperf3 and handles output logging in JSON, CSV, and plain text formats. Key command-line flags include:
\begin{itemize}
    \item \texttt{--server}: Runs the iperf3 server in JSON output mode.
    \item \texttt{<SERVER\_IP>}: Specifies the server's IP address for client mode.
    \item \texttt{--udp}: Switches the test from TCP to UDP.
    \item \texttt{--bitrate}: Sets the target bitrate for UDP tests (e.g., \texttt{10M}).
    \item \texttt{--iter}: Specifies the number of test iterations.
\end{itemize}
The output files generated by the script serve as the basis for our statistical analysis and subsequent comparison with the theoretical predictions.
